Beta Diversity through Time and Space

### 2021-04-07 code/ms1_beta_diversity.Rmd
updated analyses and figures ot include mamagement; dropped MRM analysis

### Prepare environment
```{r message=FALSE, warning=FALSE}
library(sf)
library(gdm)
library(patchwork)
library(bipartite)
library(vegan)
library(see)
library(tidyverse)
library(lubridate)
```

@@

predData_prep_survey <- function(surv) {
  surv %>%
    filter(flower.cover > 0) %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    dplyr::select(site, elev.mean, y, x, samples) %>%
    unique() %>%
    arrange(site) 
}

@@

predData_prep_BB <- function(net) {
  net %>%
    filter(caste != "male") %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    dplyr::select(site, elev.mean, y, x, samples) %>%
    unique() %>%
    arrange(site) 
}
```

@@

### MRM analysis
```{r}
library(ecodist)

survey_sample_diff <- survey %>%
  group_by(site) %>%
  summarize(samples = length(unique(date))) %>%
  remove_rownames() %>%
  column_to_rownames("site") %>%
  vegdist(method = "euclidean", upper = TRUE) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "sample.diff")

@@

mrm_fl <- MRM(beta ~ elev.diff + sample.diff + mgmt.diff, 
              data = mrm_survey_data,
              nperm = 1000,
              method = "logistic")

mrm_bb <- MRM(beta ~ elev.diff + sample.diff + mgmt.diff, 
              data = mrm_net_data,
              nperm = 1000,
              method = "logistic")

mrm_WN <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "WN"),
              nperm = 1000,
              method = "logistic")

mrm_ST <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST"),
              nperm = 1000,
              method = "logistic")

mrm_ST.l <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST.l"),
              nperm = 1000,
              method = "logistic")

mrm_ST.h <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST.h"),
              nperm = 1000,
              method = "logistic")

mrm_ST.lh <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST.lh"),
              nperm = 1000,
              method = "logistic")

mrm_OS <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "OS"),
              nperm = 1000,
              method = "logistic")

# Partial Mantel tests are another option, perhaps more routine to interpret
# mantel_fl <- mantel(beta ~ elev.diff + sample.diff, 
#               data = mrm_survey_data,
#               mrank = TRUE)
# 
# mantel_bb <- mantel(beta ~ elev.diff + sample.diff, 
#               data = mrm_net_data,
#               mrank = TRUE)
# 
# mantel_WN <- mantel(value ~ elev.diff + sample.diff, 
#                     data = filter(mrm_beta_data, metric == "WN"),
#                     mrank = TRUE)
# 
# mantel_ST <- mantel(value ~ elev.diff + sample.diff, 
#                     data = filter(mrm_beta_data, metric == "ST"),
#                     mrank = TRUE)
# 
# mantel_ST.l <- mantel(value ~ elev.diff + sample.diff, 
#                     data = filter(mrm_beta_data, metric == "ST.l"),
#                     mrank = TRUE)
# 
# mantel_ST.h <- mantel(value ~ elev.diff + sample.diff, 
#                     data = filter(mrm_beta_data, metric == "ST.h"),
#                     mrank = TRUE)
# 
# mantel_ST.lh <- mantel(value ~ elev.diff + sample.diff, 
#                     data = filter(mrm_beta_data, metric == "ST.lh"),
#                     mrank = TRUE)
# 
# mantel_OS <- mantel(value ~ elev.diff + sample.diff, 
#                     data = filter(mrm_beta_data, metric == "OS"),
#                     mrank = TRUE)
```

### GDM: Interactions
```{r message=FALSE, warning=FALSE}
# Prep sitepairs
sitepair.WN <- interaction_gdm_prep(net)

# Run GDM analysis
gdm_WN <- gdm(sitepair.WN, geo = FALSE)
summary(gdm_WN)
gdm.varImp_WN <- gdm.varImp(sitepair.WN, geo = FALSE)

@@

# Plot with confidence intervals 
pdf("../output/gdm_interactions.pdf", height = 2.5, width =  6.5)
plotUncertainty(sitepair.WN, sampleSites = 0.7, bsIters = 100, geo  = TRUE, 
                splineCol = "black", plot.layout = c(1,3))
dev.off()
```

### GDM: Floral survey
```{r}
# Prep input for GDM
survey_bioData <- biodata_prep_survey(survey)
survey_predData <- predData_prep_survey(survey)
survey_sitepair <- format_set(survey_bioData, survey_predData)

# Call GDMs
gdm_survey <- gdm(survey_sitepair, geo = TRUE)
summary(gdm_survey)

@@

# Plot with confidence intervals (slow)
pdf("../output/gdm_flora.pdf", height = 2.5, width =  6.5)
plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 100, geo  = TRUE, 
                splineCol = "#00BFC4", plot.layout = c(1,3))
dev.off()
```

### GDM: BB observations
```{r}
# Prep input for GDM
BB_bioData <- biodata_prep_BB(net)
BB_predData <- predData_prep_BB(net)
BB_sitepair <- format_set(BB_bioData, BB_predData)

# Call GDMs
gdm_BB <- gdm(BB_sitepair, geo = TRUE)
summary(gdm_BB)

@@

# Plot with confidence intervals (slow)
pdf("../output/gdm_BB.pdf", height = 2.5, width =  6.5)
plotUncertainty(BB_sitepair, sampleSites = 0.7, bsIters = 100, geo  = TRUE, 
                splineCol = "#F8766D", plot.layout = c(1,3))
dev.off()
```

### Tabulate model output
```{r}
# assign("gdm_WN.2010.imp", get(load("~/Documents/Research/bb_berchtesgaden/analyses/output/interaction_permGDM_2010.RData")))
# 
# assign("gdm_WN.2011.imp", get(load("~/Documents/Research/bb_berchtesgaden/analyses/output/interaction_permGDM_2011.RData")))
# 
# assign("gdm_WN.2012.imp", get(load("~/Documents/Research/bb_berchtesgaden/analyses/output/interaction_permGDM_2012.RData")))

####### WN
gdm_WN.2010.tab1 <- gdm_WN.2010.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "WN",
         year = 2010) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_WN.2010.tab2 <- gdm_WN.2010.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_WN.2010.tab3 <- gdm_WN.2010.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_WN.2011.tab1 <- gdm_WN.2011.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "WN",
         year = 2011) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_WN.2011.tab2 <- gdm_WN.2011.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_WN.2011.tab3 <- gdm_WN.2011.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_WN.2012.tab1 <- gdm_WN.2012.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "WN",
         year = 2012) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_WN.2012.tab2 <- gdm_WN.2012.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_WN.2012.tab3 <- gdm_WN.2012.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_WN.2010.tab_join <- bind_cols(gdm_WN.2010.tab1,
                                  gdm_WN.2010.tab2, 
                                  gdm_WN.2010.tab3)

gdm_WN.2011.tab_join <- bind_cols(gdm_WN.2011.tab1,
                                  gdm_WN.2011.tab2, 
                                  gdm_WN.2011.tab3)

gdm_WN.2012.tab_join <- bind_cols(gdm_WN.2012.tab1,
                                  gdm_WN.2012.tab2, 
                                  gdm_WN.2012.tab3)

####### survey
gdm_survey.2010.tab1 <- gdm_survey.2010.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "survey",
         year = 2010) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_survey.2010.tab2 <- gdm_survey.2010.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_survey.2010.tab3 <- gdm_survey.2010.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_survey.2011.tab1 <- gdm_survey.2011.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "survey",
         year = 2011) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_survey.2011.tab2 <- gdm_survey.2011.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_survey.2011.tab3 <- gdm_survey.2011.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_survey.2012.tab1 <- gdm_survey.2012.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplylr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "survey",
         year = 2012) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_survey.2012.tab2 <- gdm_survey.2012.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_survey.2012.tab3 <- gdm_survey.2012.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplytr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_survey.2010.tab_join <- bind_cols(gdm_survey.2010.tab1,
                                  gdm_survey.2010.tab2, 
                                  gdm_survey.2010.tab3)

gdm_survey.2011.tab_join <- bind_cols(gdm_survey.2011.tab1,
                                  gdm_survey.2011.tab2, 
                                  gdm_survey.2011.tab3)

gdm_survey.2012.tab_join <- bind_cols(gdm_survey.2012.tab1,
                                  gdm_survey.2012.tab2, 
                                  gdm_survey.2012.tab3)

######## BB
gdm_BB.2010.tab1 <- gdm_BB.2010.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "BB",
         year = 2010) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_BB.2010.tab2 <- gdm_BB.2010.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_BB.2010.tab3 <- gdm_BB.2010.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_BB.2011.tab1 <- gdm_BB.2011.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "BB",
         year = 2011) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_BB.2011.tab2 <- gdm_BB.2011.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_BB.2011.tab3 <- gdm_BB.2011.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_BB.2012.tab1 <- gdm_BB.2012.imp[1] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  mutate(subject = "BB",
         year = 2012) %>%
  dplyr::select(subject, year, 
         mod.dev = `Model deviance`, 
         dev.exp = `Percent deviance explained`, 
         mod.pval = `Model p-value`)

gdm_BB.2012.tab2 <- gdm_BB.2012.imp[2] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble() %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2) %>%
  dplyr::select(geo.imp = Geographic, elev.imp = elev.mean, yday.imp = yday)

gdm_BB.2012.tab3 <- gdm_BB.2012.imp[3] %>% 
  as.data.frame() %>%
  rownames_to_column() %>%
  as_tibble()  %>%
  dplyr::select(1, 2) %>%
  pivot_wider(names_from = rowname, values_from = 2)  %>%
  dplyr::select(geo.pval = Geographic, elev.pval = elev.mean, yday.pval = yday)

gdm_BB.2010.tab_join <- bind_cols(gdm_BB.2010.tab1,
                                  gdm_BB.2010.tab2, 
                                  gdm_BB.2010.tab3)

gdm_BB.2011.tab_join <- bind_cols(gdm_BB.2011.tab1,
                                  gdm_BB.2011.tab2, 
                                  gdm_BB.2011.tab3)

gdm_BB.2012.tab_join <- bind_cols(gdm_BB.2012.tab1,
                                  gdm_BB.2012.tab2, 
                                  gdm_BB.2012.tab3)

#######################

gdm_BB.tab_join <- bind_rows(gdm_BB.2010.tab_join,
                             gdm_BB.2011.tab_join,
                             gdm_BB.2012.tab_join,
                             gdm_survey.2010.tab_join,
                             gdm_survey.2011.tab_join,
                             gdm_survey.2012.tab_join,
                             gdm_WN.2010.tab_join,
                             gdm_WN.2011.tab_join,
                             gdm_WN.2012.tab_join) %>%
  dplyr::select(-subject)

gdm_tab <- gdm_BB.tab_join %>%
  gt(rowname_col = "year") %>%
  tab_spanner(
    label = "Model",
    columns = vars(mod.dev, dev.exp, mod.pval)
  ) %>%
  tab_spanner(
    label = "Geographic distance",
    columns = vars(geo.imp, geo.pval)
  ) %>%
  tab_spanner(
    label = "Elevational distance",
    columns = vars(elev.imp, elev.pval)
  ) %>%
  tab_spanner(
    label = "Temporal distance",
    columns = vars(yday.imp, yday.pval)
  ) %>%
  tab_row_group(
    group = "Bumble bees",
    rows = 1:3
  ) %>% 
  tab_row_group(
    group = "Flora",
    rows = 4:6
  ) %>%
  tab_row_group(
    group = "Interactions",
    rows = 7:9
  ) %>%
  fmt_number(
    columns = c("mod.dev", "dev.exp",
                "geo.imp", "elev.imp", "yday.imp"),
    decimals = 1
  ) %>%
  fmt_number(
    columns = c("mod.pval", "geo.pval", 
                "elev.pval", "yday.pval"),
    decimals = 3
  ) %>%
  cols_label(mod.dev = "tot. dev.",
             dev.exp = "dev. expl. (%)",
             mod.pval = "p",
             geo.imp = "imp.",
             elev.imp = "imp.",
             yday.imp = "imp.",
             geo.pval = "p-val",
             elev.pval = "p-val",
             yday.pval = "p-val")

>>>>

### Prepare environment
```{r message=FALSE, warning=FALSE}
library(sf)
library(gdm)
library(patchwork)
library(bipartite)
library(vegan)
library(ecodist)
library(gt)
library(tidyverse)
library(lubridate)
```

@@

predData_prep_survey <- function(surv) {
  surv %>%
    filter(flower.cover > 0) %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    mutate(mowing = if_else(management == "mowing", 1, 0),
           grasing = if_else(management == "grasing", 1, 0),
           none = if_else(management == "none", 1, 0)) %>%
    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
    distinct() %>%
    arrange(site) 
}

@@

predData_prep_BB <- function(net) {
  net %>%
    filter(caste != "male") %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    mutate(mowing = if_else(management == "mowing", 1, 0),
           grasing = if_else(management == "grasing", 1, 0),
           none = if_else(management == "none", 1, 0)) %>%
    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
    unique() %>%
    arrange(site) 
}
```

@@

### MRM analysis
```{r, warning=FALSE}
survey_sample_diff <- survey %>%
  group_by(site) %>%
  summarize(samples = length(unique(date))) %>%
  remove_rownames() %>%
  column_to_rownames("site") %>%
  vegdist(method = "euclidean", upper = TRUE) %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "sample.diff")

@@

# Function for tabulating modle coefficients
mrm_tab <- function(x, response) {
  x %>%
    as.data.frame() %>%
    rownames_to_column(var = "predictor") %>%
    slice(-1) %>%
    mutate(response = rep(response, n()),
           Predictor = case_when(
             predictor == "elev.diff" ~ "elevation",
             predictor == "sample.diff" ~ "sampling freq.",
             predictor == "mgmt.diff" ~ "management")) %>%
    select(response, Predictor, b, pval) %>%
    tibble()
}

# Call models
mrm_fl <- MRM(beta ~ elev.diff + sample.diff + mgmt.diff, 
              data = mrm_survey_data,
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("Flora")

mrm_bb <- MRM(beta ~ elev.diff + sample.diff + mgmt.diff, 
              data = mrm_net_data,
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("Bumble bees")

mrm_WN <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "WN"),
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("Total interactions")

mrm_ST <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST"),
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("ST partition")

mrm_ST.l <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST.l"),
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("ST.l partition")

mrm_ST.h <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST.h"),
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("ST.h partition")

mrm_ST.lh <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "ST.lh"),
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("ST.lh partition")

mrm_OS <- MRM(value ~ elev.diff + sample.diff + mgmt.diff, 
              data = filter(mrm_beta_data, metric == "OS"),
              nperm = 1000,
              method = "logistic")$coef %>%
  mrm_tab("Rewiring partition")

# Tabulate results
mrm_results <- bind_rows(mrm_fl, mrm_bb, mrm_WN, mrm_ST, 
                         mrm_ST.h, mrm_ST.l, mrm_ST.lh, mrm_OS) %>%
  gt(groupname_col = "response") %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = list(cells_column_labels(everything()))
  ) %>%
  tab_style(
    style = list(cell_text(style = "italic")),
    locations = list(cells_group(everything()))
  ) %>%
  # data_color(
  #   columns = vars(b),
  #   colors = scales::col_numeric(
  #     palette = c("white", "red"),
  #     domain = c(-0.01, 0.5)), alpha = 0.75
  #   ) %>%
  tab_style(
    style = list(
      #cell_fill(color = "lightcyan"),
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = vars(pval),
      rows = pval < 0.05
    ))

gtsave(mrm_results,"../output/mrm_tab.pdf")

mrm_results_sub <- bind_rows(mrm_fl, mrm_bb, mrm_WN) %>%
  gt(groupname_col = "response") %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = list(cells_column_labels(everything()))
  ) %>%
  tab_style(
    style = list(cell_text(style = "italic")),
    locations = list(cells_group(everything()))
  ) %>%
  # data_color(
  #   columns = vars(b),
  #   colors = scales::col_numeric(
  #     palette = c("white", "red"),
  #     domain = c(-0.01, 0.5)), alpha = 0.75
  #   ) %>%
  tab_style(
    style = list(
      #cell_fill(color = "lightcyan"),
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = vars(pval),
      rows = pval < 0.05
    ))
```

### GDM: Interactions
```{r message=FALSE, warning=FALSE}
# Prep sitepairs
sitepair_WN <- interaction_gdm_prep(net)

# Run GDM analysis
gdm_WN <- gdm(sitepair_WN, geo = FALSE)
summary(gdm_WN)
gdm.varImp_WN <- gdm.varImp(sitepair_WN, geo = FALSE)

@@

# Plot with confidence intervals 
# pdf("../output/gdm_interactions.pdf", height = 2.5, width =  6.5)
# plotUncertainty(sitepair.WN, sampleSites = 0.7, bsIters = 100, geo  = FALSE, 
#                 splineCol = "black", plot.layout = c(1,3))
# dev.off()
```

### GDM: Floral survey
```{r}
# Prep input for GDM
bioData_survey <- biodata_prep_survey(survey)
predData_survey <- predData_prep_survey(survey)
sitepair_survey <- format_set(bioData_survey, predData_survey)

# Call GDMs
gdm_survey <- gdm(survey_sitepair, geo = FALSE)
summary(gdm_survey)
gdm.varImp_survey <- gdm.varImp(sitepair_survey, geo = FALSE)

@@

# Plot with confidence intervals (slow)
# pdf("../output/gdm_flora.pdf", height = 2.5, width =  6.5)
# plotUncertainty(sitepair_survey, sampleSites = 0.7, bsIters = 100, geo  = FALSE, 
#                 splineCol = "#00BFC4", plot.layout = c(1,3))
# dev.off()
```

### GDM: BB observations
```{r}
# Prep input for GDM
bioData_BB <- biodata_prep_BB(net)
predData_BB <- predData_prep_BB(net)
sitepair_BB <- format_set(bioData_BB, predData_BB)

# Call GDMs
gdm_BB <- gdm(sitepair_BB, geo = FALSE)
summary(gdm_BB)
gdm.varImp_BB <- gdm.varImp(sitepair_BB, geo = FALSE)

@@

# Plot with confidence intervals (slow)
# pdf("../output/gdm_BB.pdf", height = 2.5, width =  6.5)
# plotUncertainty(sitepair_BB, sampleSites = 0.7, bsIters = 100, geo  = TRUE, 
#                 splineCol = "#F8766D", plot.layout = c(1,3))
# dev.off()
```

### Tabulate gdm.varImp results
```{r}
# Function for tabulate gdm.varImp output 
gdm_tab <- function(x, response) {
  x %>%
    as.data.frame() %>%
    rownames_to_column(var = "predictor") %>%
    mutate(response = rep(response, n()),
           Predictor = case_when(
             predictor == "elev.mean" ~ "elevation",
             predictor == "samples" ~ "sampling freq.",
             predictor == "mowing" ~ "mowing",
             predictor == "grasing" ~ "grasing",
             predictor == "none" ~ "none")) %>%
    select(response, Predictor, Importance, pval) %>%
    tibble()
}

# Interaction model
gdm.varImp_WN.tab2 <- gdm.varImp_WN[[2]] %>%
  as.data.frame() %>%
  select(Importance = 1)

gdm.varImp_WN.tab3 <- gdm.varImp_WN[[3]] %>%
  as.data.frame() %>%
  select(pval = 1)

gdm.varImp_WN.tab <- bind_cols(gdm.varImp_WN.tab2, gdm.varImp_WN.tab3) %>%
  gdm_tab("Interactions")

# Floral model
gdm.varImp_survey.tab2 <- gdm.varImp_survey[[2]] %>%
  as.data.frame() %>%
  select(Importance = 1)

gdm.varImp_survey.tab3 <- gdm.varImp_survey[[3]] %>%
  as.data.frame() %>%
  select(pval = 1)

gdm.varImp_survey.tab <- bind_cols(gdm.varImp_survey.tab2, gdm.varImp_survey.tab3) %>%
  gdm_tab("Flora")

# BB model
gdm.varImp_BB.tab2 <- gdm.varImp_BB[[2]] %>%
  as.data.frame() %>%
  select(Importance = 1)

gdm.varImp_BB.tab3 <- gdm.varImp_BB[[3]] %>%
  as.data.frame() %>%
  select(pval = 1)

gdm.varImp_BB.tab <- bind_cols(gdm.varImp_BB.tab2, gdm.varImp_BB.tab3) %>%
  gdm_tab("Bumble bees")

# Merge and format
gdm.varImp.tab <- bind_rows(gdm.varImp_WN.tab, gdm.varImp_survey.tab, gdm.varImp_BB.tab) %>%
  mutate(Importance = round(Importance, 2)) %>%
  gt(groupname_col = "response") %>%
  tab_style(
    style = list(cell_text(weight = "bold")),
    locations = list(cells_column_labels(everything()))
  ) %>%
  tab_style(
    style = list(cell_text(style = "italic")),
    locations = list(cells_row_groups(everything()))
  ) %>%
  data_color(
    columns = vars(Importance),
    colors = scales::col_numeric(
      palette = c("white", "red"),
      domain = c(-0, 56)), alpha = 0.75
    ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = vars(pval),
      rows = pval < 0.05
    ))

gtsave(gdm.varImp.tab, "../output/gdm_tab.pdf")
```

### Plot GDM splines
```{r}
# Interaction model
gdm_WN_splines1 <- isplineExtract(gdm_WN)$x %>%
  as.data.frame() %>%
  select(elevation = elev.mean)

gdm_WN_splines2 <- isplineExtract(gdm_WN)$y %>%
  as.data.frame() %>%
  select(response = elev.mean)

gdm_WN_splines <- bind_cols(gdm_WN_splines1, gdm_WN_splines2) %>%
  mutate(level = rep("Interactions"))

# Flora model
gdm_survey_splines1 <- isplineExtract(gdm_survey)$x %>%
  as.data.frame() %>%
  select(elevation = elev.mean)

gdm_survey_splines2 <- isplineExtract(gdm_survey)$y %>%
  as.data.frame() %>%
  select(response = elev.mean)

gdm_survey_splines <- bind_cols(gdm_survey_splines1, gdm_survey_splines2) %>%
  mutate(level = rep("Flora"))

# BB model
gdm_BB_splines1 <- isplineExtract(gdm_BB)$x %>%
  as.data.frame() %>%
  select(elevation = elev.mean)

gdm_BB_splines2 <- isplineExtract(gdm_BB)$y %>%
  as.data.frame() %>%
  select(response = elev.mean)

gdm_BB_splines <- bind_cols(gdm_BB_splines1, gdm_BB_splines2) %>%
  mutate(level = rep("Bumble bees"))

# Combine
gdm_splines <- bind_rows(gdm_WN_splines, gdm_survey_splines, gdm_BB_splines) %>%
  mutate(Level = factor(level, levels = c("Interactions", "Flora", "Bumble bees")))

tree_line <- site_data %>%
  select(elevation = elev.mean, tree_line)

# Plot
ggplot(gdm_splines, aes(elevation, response, linetype = Level)) +
  geom_line() +
  #geom_vline(xintercept = 1500) +
  geom_rug(data = tree_line, aes(elevation, color = tree_line), inherit.aes = FALSE) +
  labs(color = "Tree Line", x = "Elevation", y = "Response") +
  theme_light(14)

ggsave("../output/gdm_splines.pdf")
ggsave("../output/gdm_splines.png")
